---
title: "Unidentified Aerial Phenomenon in America"
author: Tev'ye davish
subtitle: 
---

# Introduction

[\~ 200 words]

An unidentified flying object (UFO), recently renamed as a UAP (unidentified aerial phenomenon) are observations of events in the sky that cannot be identified as aircraft or as known natural phenomena. Understanding the data surrounding unidentified aerial phenomena is critical to helping us draw scientific conclusions about what is happening in the sky. Unidentified aerial phenomena are of interest, from a security and safety perspective. There have been several studies of UAPs carried out by various US government agencies, including a Pentagon report that was declassified in June 2021. The US House of Representatives recently voted to encourage the sharing of more UFO sightings by adopting a bipartisan amendment to the National Defense Authorization Act. The aim of this study is to create a map depicting the number of sightings of UAP per state and identify if there is a cluster of UAP sightings.

# Materials and methods

[\~ 200 words]

Point locations of UAP/UFO sightings were from the National UFO Reporting Center. The National UFO Reporting Center was founded in 1974 and has documented approximately 90,000 reported UFO sightings over its history mostly in the United States. The shape file for the states will be obtained from the tigris package.

Here's my first code chunk.

Load any required packages in a code chunk (you may need to install some packages):

```{r, message=F, warning=F}
library(tidyverse)
library(leaflet)
library(htmltools)
#install.packages("tidycensus")
#library(tidycensus)
#install.packages("tigris")
library(tigris)
library(maptools)
library(spdep)
library(sf)
library(lubridate)
library(viridis)
knitr::opts_chunk$set(cache=TRUE)  # cache the results for quick compiling
```

## Download and clean all required data

```{r, message = FALSE, warning = FALSE}
uap = read_csv('https://query.data.world/s/6xru2x6vjkjgz5ck52invxt3j7mc5w')

US_map = states(cb = T)
US_map = as(US_map, 'sf')
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r, message = FALSE, warning = FALSE}
US_map = states(cb = T)
US_map = as(US_map, 'sf')
```

```{r, results='asis'}
us_states = c('AL', 'AZ', 'AR', 'CA', 'CO', 'CT', 'DE', 'FL',
              'GA', 'ID', 'IL', 'IN', 'IA', 'KS', 'KY', 'LA',
              'ME', 'MD', 'MA', 'MI', 'MN', 'MS', 'MO', 'MT', 'NE',
              'NV', 'NH', 'NJ', 'NM', 'NY', 'NC', 'ND', 'OH', 'OK',
              'OR', 'PA', 'RI', 'SC', 'SD', 'TN', 'TX', 'UT', 'VT',
              'VA', 'WA', 'WV', 'WI', 'WY', 'DC')

US_map = US_map %>% filter(STUSPS %in% us_states)


```

Add any additional processing steps here.

# Results

[\~200 words]

Tables and figures (maps and other graphics) are carefully planned to convey the results of your analysis. Intense exploration and evidence of many trials and failures. The author looked at the data in many different ways before coming to the final presentation of the data.

Show tables, plots, etc. and describe them.

```{r, fig.width=8, fig.height=5, fig.cap="Number of UAPs per State"}
US_uap = uap %>% filter(state %in% us_states) %>% group_by(state) %>% summarize(n = n()) 

US_map = left_join(US_map, US_uap, by = c("STUSPS" = 'state'))

US_map %>% 
  ggplot() +
  geom_col(aes(x = fct_reorder(STUSPS, n), y = n)) + coord_flip() +
  theme_light() +
  labs(x = 'Number of UAPs', y = 'State') +
  geom_text(aes(x = fct_reorder(STUSPS, n), y = n, label = n), nudge_y = 350, size = 3.5) +
  ggtitle('Number of UAPs per State')
```

```{r, message = FALSE, warning = FALSE}
uap$year = year(uap$date_time)
uap %>% select(year, state) %>%  na.omit() %>% 
  group_by(year, state) %>% 
  summarize(n = n()) %>% ungroup() %>% 
  group_by(year) %>% summarize(tot_uaps = sum(n)) %>% 
  ggplot() +
  geom_line(aes(x = year, y = tot_uaps)) +
  geom_point(aes(x = year, y = tot_uaps)) +
  theme_light() +
  labs(x = 'Year', y = 'UAP sightings') + 
  ggtitle('UAP sightings in the USA from 1969 to 2022') +
  scale_x_continuous(breaks = seq(1969, 2022, 5))
```

```{r, message = FALSE, warning = FALSE}
uap %>% select(year, state) %>%  na.omit() %>% 
  filter(state %in% us_states[1:10]) %>% 
  group_by(year, state) %>% 
  summarize(n = n()) %>% ungroup() %>% 
  group_by(year, state) %>% summarize(tot_uaps = sum(n)) %>% 
  ggplot() +
  geom_line(aes(x = year, y = tot_uaps, color = state)) +
  geom_point(aes(x = year, y = tot_uaps, color = state)) +
  theme_light() +
  labs(x = 'Year', y = 'UAP sightings') + 
  ggtitle('UAP sightings in the USA from 1969 to 2022') +
  scale_x_continuous(breaks = seq(1969, 2022, 5)) +
  scale_color_discrete(name = 'State') + ylim(0, 900)
```

```{r, message = FALSE, warning = FALSE}
uap %>% select(year, state) %>%  na.omit() %>% 
  filter(state %in% us_states[11:20]) %>% 
  group_by(year, state) %>% 
  summarize(n = n()) %>% ungroup() %>% 
  group_by(year, state) %>% summarize(tot_uaps = sum(n)) %>% 
  ggplot() +
  geom_line(aes(x = year, y = tot_uaps, color = state)) +
  geom_point(aes(x = year, y = tot_uaps, color = state)) +
  theme_light() +
  labs(x = 'Year', y = 'UAP sightings') + 
  ggtitle('UAP sightings in the USA from 1969 to 2022') +
  scale_x_continuous(breaks = seq(1969, 2022, 5)) + 
  scale_color_discrete(name = 'State') + ylim(0, 900)
```

```{r, message = FALSE, warning = FALSE}
uap %>% select(year, state) %>%  na.omit() %>% 
  filter(state %in% us_states[21:30]) %>% 
  group_by(year, state) %>% 
  summarize(n = n()) %>% ungroup() %>% 
  group_by(year, state) %>% summarize(tot_uaps = sum(n)) %>% 
  ggplot() +
  geom_line(aes(x = year, y = tot_uaps, color = state)) +
  geom_point(aes(x = year, y = tot_uaps, color = state)) +
  theme_light() +
  labs(x = 'Year', y = 'UAP sightings') + 
  ggtitle('UAP sightings in the USA from 1969 to 2022') +
  scale_x_continuous(breaks = seq(1969, 2022, 5)) + 
  scale_color_discrete(name = 'State') + ylim(0, 900)
```

```{r, message = FALSE, warning = FALSE}
uap %>% select(year, state) %>%  na.omit() %>% 
  filter(state %in% us_states[31:40]) %>% 
  group_by(year, state) %>% 
  summarize(n = n()) %>% ungroup() %>% 
  group_by(year, state) %>% summarize(tot_uaps = sum(n)) %>% 
  ggplot() +
  geom_line(aes(x = year, y = tot_uaps, color = state)) +
  geom_point(aes(x = year, y = tot_uaps, color = state)) +
  theme_light() +
  labs(x = 'Year', y = 'UAP sightings') + 
  ggtitle('UAP sightings in the USA from 1969 to 2022') +
  scale_x_continuous(breaks = seq(1969, 2022, 5)) + 
  scale_color_discrete(name = 'State') + ylim(0, 900)
```

```{r, message = FALSE, warning = FALSE}
uap %>% select(year, state) %>%  na.omit() %>% 
  filter(state %in% us_states[41:49]) %>% 
  group_by(year, state) %>% 
  summarize(n = n()) %>% ungroup() %>% 
  group_by(year, state) %>% summarize(tot_uaps = sum(n)) %>% 
  ggplot() +
  geom_line(aes(x = year, y = tot_uaps, color = state)) +
  geom_point(aes(x = year, y = tot_uaps, color = state)) +
  theme_light() +
  labs(x = 'Year', y = 'UAP sightings') + 
  ggtitle('UAP sightings in the USA from 1969 to 2022') +
  scale_x_continuous(breaks = seq(1969, 2022, 5)) + 
  scale_color_discrete(name = 'State') + ylim(0, 900)
```
```{r, message = FALSE, warning = FALSE}
US_map_nb = poly2nb(US_map)
US_map_listw = nb2listw(US_map_nb)

US_coords = coordinates(as_Spatial(US_map)) %>% data.frame()

neighborhood = as(spdep::nb2lines(US_map_nb, coords = coordinates(as_Spatial(US_map))), 'sf')
neighborhood = sf::st_set_crs(neighborhood, sf::st_crs(US_map))

ggplot(US_map) +
  geom_sf(size = 1, color = 'gray', fill = 'white') +
  geom_sf(data = neighborhood) +
  geom_point(data = US_coords, aes(x = X1, y = X2), color = 'red') +
  theme_light() +
  ggtitle('Contiguous US neighborhood structure')
```

```{r, message = FALSE, warning = FALSE}
labels <- sprintf(
  "<strong>%s</strong><br/>%g uaps",
  US_map$NAME, US_map$n
) %>% lapply(htmltools::HTML)

pal = colorBin('YlOrRd', domain = US_map$n, bins = quantile(US_map$n, c(seq(0,0.9, by = 0.1125), 1)))

leaflet(US_map) %>% 
  addPolygons(
    fillColor = ~pal(n),
    fillOpacity = 1, 
    color = 'white',
    weight = 2,
    label = labels,
    highlight = highlightOptions(
      weight = 4,
      color = 'black',
      fillOpacity = 1,
      bringToFront = TRUE)) %>%
  addProviderTiles(providers$Esri.WorldStreetMap) %>% 
  addLegend(pal = pal, values = ~n, opacity = 0.7, title = 'UAP Sightings: 1969 - 2022',
            position = "bottomright")
```

```{r, message = FALSE, warning = FALSE}
moran.mc(US_map$n, listw = US_map_listw, nsim = 9999)

```

# Conclusions

[\~200 words]

Clear summary adequately describing the results and putting them in context. Discussion of further questions and ways to continue investigation.

# References

All sources are cited in a consistent manner
